Covariate analysis of ROSMAP RNASeq

### 2016-09-03 code/Rmd/ROSMAP_RNASeq.Rmd
Minor modifications

## Load required libraries
library(synapseClient)
library(ggplot2)
library(reshape2)
library(data.table)
library(dplyr)
library(stringr)
library(limma)
library(knitr)
library(Biobase)
library(GEOquery)
library(RColorBrewer)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)
library(ggplot2)
library(WGCNA)

## Needs the dev branch
library(rGithubClient)

@@

# source utility files from ../R/lib folder
file.sources = list.files('../R/lib',pattern="*.R", full.names=T)
tmp = sapply(file.sources,source,.GlobalEnv)

@@

# Match technical and clinical covariates
METADATA <- dplyr::inner_join(METADATA.CLINICAL,METADATA.TECH)
```

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  mutate(Sampleid = paste(Sampleid, Batch, sep = '_')) %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi)) %>%
  filter(!(Sampleid %in% c('380_120503_1')))
```
Sample id 380\_120503\_1 is considered as outlier and removed
```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))

@@

@@

@@

COVARIATES.tmp = data.matrix(COVARIATES)
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
```
```{r decompse.normalise.data4, fig.height=8, fig.width=16, results='asis'}
tree = hclust(as.dist(t(RESIDUAL.GENE_EXPRESSION[,-(1:2)])))
WGCNA::plotDendroAndColors(tree, 
                    colors = labels2colors(COVARIATES.tmp), 
                    dendroLabels = FALSE, 
                    abHeight = 0.80, 
                    main = "Sample dendrogram",
                    groupLabels = colnames(COVARIATES.tmp))
dev.off()
```

>>>>

## Load required libraries
library(CovariateAnalysis) # get it from devtools::install_github('th1vairam/CovariateAnalysis@dev')
library(data.table)
library(tidyr)
library(plyr)
library(dplyr)
library(stringr)

library(ggplot2)
library(reshape2)
library(limma)
library(Biobase)
library(gplots)
library(WGCNA)
library(psych)
library(edgeR)
library(biomaRt)

@@

@@

# Fix error in technical covariates data
KEY_ID <- 'syn3382527'
ALL_USED_IDs[length(ALL_USED_IDs)+1] = KEY_ID  
KEY <- read.csv(synGet(KEY_ID)@filePath) %>% 
  filter(mrna_data == 1) %>%
  dplyr::select(projid, mrna_id) %>%
  separate(mrna_id, c('a','b','batch'), sep = '_') %>% 
  unite(Sampleid, a, b) %>%
  dplyr::select(-batch) %>%
  unique

# Match technical and clinical covariates
METADATA <- METADATA.TECH %>% 
  dplyr::select(-projid) %>%
  dplyr::left_join(KEY) %>%
  dplyr::left_join(METADATA.CLINICAL)

# Pick higher quality RIN batch for sample 492_120515
METADATA <- METADATA %>%
  group_by(Sampleid) %>%
  top_n(1, RINcontinuous)
```

### Data preprocessing
```{r preprocess.data,cache=TRUE, echo=TRUE}
# Remove samples with no cogdx, RIN and PMI scores
METADATA <- METADATA %>%
  ungroup %>%
  mutate(Sampleid = paste(Sampleid, Batch, sep = '_')) %>%
  filter(Sampleid %in% colnames(COUNT)) %>%
  filter(!is.na(cogdx)) %>%
  filter(!is.na(RINcontinuous)) %>%
  filter(!is.na(pmi)) %>%
  filter(!(Sampleid %in% c('380_120503_1'))) %>%
  as.data.frame()
```
Sample id 380\_120503\_1 is considered as outlier and removed
```{r preprocess.data1,cache=TRUE}
indToRetain = intersect(METADATA$Sampleid, colnames(COUNT))

@@

Coexpression of genes 
```{r coexp1, cache=FALSE}
cr = cor(t(EXPR))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

Coexpression of genes 
```{r coexp2, cache=FALSE}
cr = cor(t(RESIDUAL.GENE_EXPRESSION[, !(colnames(RESIDUAL.GENE_EXPRESSION) %in% c('ensembl_gene_id','position'))]))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```

@@

COVARIATES.tmp = data.matrix(COVARIATES)
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0
```
